---
title: "Thermocline computation"
author: "Vilem Ded"
date: "9/28/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = F, warning = F)

```

```{r}
library(here)
source(here("R", "00_config.R"))
source(here("R", "99_functions.R"))
```

## Thermocline Computation
This document shows how was the thermocline measured, computed and interpolated into the whole lake. It also shows some issues found and describes the accepted solutions.
```{r}
# Loading all data
temperatures <- load_temperature_data() %>% as_tibble

logger_positions <- load_logger_positions()

hobo_deployments <- read_csv(here("data", "raw", "db", "hobodeployements.csv")) %>%
  filter(month(retrieval_ts) == 8)

hobo_data <- as_tibble(load_hobo_data())

lake_shp <- st_read(here("data/raw/db/lake_shape.shp"), quiet = T)
lake_axis <- st_read(here("data/raw/sp/lake_axis.shp"), quiet = T)
# get shifted axis into deep part
pts <- data.frame(name = "blue", x = 40, y = 120)
pts <- st_as_sf(pts, coords = c('x', 'y'), crs = 32633)
lake_axis_shifted <- lake_axis
lake_axis_shifted$geometry <- lake_axis_shifted$geometry + pts$geometry
st_crs(lake_axis_shifted$geometry) <- st_crs(lake_axis$geometry)

hydrophones <- st_read(here("data/raw/db/hydrophones.shp"), quiet = T)

point_zero <- st_read(here("data/raw/sp/point_zero.shp"), quiet = T)

logger_positions$dist_from_zero <- compute_distance_from_point_zero(positions = logger_positions, lake_axis = lake_axis, point_zero = point_zero)


lake_raster <- raster::raster(here("data/raw/db/lake_raster.tif"))


  
```



## Loggers and data
lakeshape, logger positions, logger depths, measured data interpolation into depth

#### Lake plan view
```{r}
# categorize depths
lake_raster_df <- as_tibble(rasterToPoints(lake_raster, spatial = TRUE)) %>% 
  dplyr::rename(depth = lake_raster) %>%
  mutate(depth_cat = cut(depth, breaks = c(-Inf, 10, 15, 20, Inf), labels = c("0-5", "5-10", "10-20", ">20")))

# Get raster axis profile
axis_points <- st_line_sample(lake_axis_shifted, density = 1, type = "regular") %>% 
  st_cast("POINT") %>%
  as("Spatial")

axis_points$distance <- (1:length(axis_points))-1
axis_depth_profile <- extract(lake_raster, axis_points, sp = T, fun = mean)

hydrophones_loggers <- rbind(hydrophones %>% mutate(device = "hydrophone") %>% dplyr::select(device, geometry), 
                             logger_positions %>% mutate(device = "logger") %>% dplyr::select(device, geometry))
gp_lake_plan_view <- ggplot() + 
  geom_sf(data = lake_shp) +
  geom_raster(data = lake_raster_df, aes(x = x, y = y, fill = depth_cat)) +
  scale_fill_manual(values = c("0-5" = "#E8E8E8", "5-10" = "#C0C0C0", "10-20" = "#808080", ">20" = "#484848")) +
  geom_sf(data = hydrophones_loggers, aes(pch = device, size = device), fill = "grey") +
  scale_shape_manual(values = c("logger" = 23, "hydrophone" = 16), labels = c("logger" = "Loggers", "hydrophone" = "Hydrophones")) +
  scale_size_manual(values = c("logger" = 3, "hydrophone" = 2), guide = NULL) +
  geom_sf(data = lake_axis_shifted, aes(lty = "Lake axis")) + 
  scale_linetype_manual(values = c("Lake axis" = 2)) +
  annotation_scale(location = "bl", width_hint = 0.5) +
  annotation_north_arrow(location = "bl",
                         which_north = "true",
                         pad_x = unit(0.75, "in"), pad_y = unit(0.5, "in"),
                         style = north_arrow_fancy_orienteering) +
  xlab("Longitude") +
  ylab("Latitude") +
  labs(fill = "Depth (m)", 
       shape = "Devices",
       linetype = "")

gp_lake_plan_view 
```

#### Lake side view
```{r}
#Loggers
hobo_deploymets_sp <- hobo_deployments %>% 
  rbind(hobo_deployments %>% filter(deployment_depth == 5.4 & location == "West") %>%
  mutate(location = "East")) %>%
  inner_join(logger_positions, by = "location")

# "2015-07-08 18:55:00" 
# "2015-09-06 10:00:00"
# 2015-07-25 19:20:00
profile_x <- temperatures %>% 
  filter(ts ==  "2015-09-06 10:00:00" & (round(depth,1) * 10) %% 2 == 0 ) %>%
  inner_join(logger_positions, by = "location") %>%
  group_nest(depth) %>%
  mutate(h_interpolated_profile = map(.x = data, .f = function(x){
    x_new <- seq(0, 3214, 10)
    y_new <- interpolate_thermocline_value_linear(logger_distances = x$dist_from_zero,
                                         logger_values = x$temperature,
                                         detection_distances = x_new
                                         )
    tibble(dist_from_zero = x_new,
           temperature = y_new)
  })) %>% 
  dplyr::select(-data) %>%
  unnest(h_interpolated_profile)

jet.colors <- colorRampPalette(c("#00007F", "blue", "#007FFF", "cyan", "#7FFF7F", "yellow", "#FF7F00", "red", "#7F0000"))

axis_depth_profile_data <-  rbind(axis_depth_profile@data,
                                   tibble(distance = c(-1, max(axis_depth_profile@data$distance)+1), lake_raster = 0))
   
gp_lake_side_view <- ggplot() +
  geom_tile(data = profile_x %>% filter( dist_from_zero > 0 & dist_from_zero < 3210), 
            aes(x = dist_from_zero, y = depth, fill = temperature)) +
  scale_fill_gradientn(colours=jet.colors(7)) +
  geom_ribbon(data =axis_depth_profile_data,
                aes(x = distance, ymin = max(lake_raster, na.rm = T), ymax = lake_raster), col = "black") +
  geom_path(data = 
              hobo_deploymets_sp %>% group_by(location, dist_from_zero) %>% summarise(depth = range(deployment_depth), .groups = "keep"), 
            aes(x = dist_from_zero, y = depth, group = location, col = " ")) +
  geom_point(data = hobo_deploymets_sp, aes(x = dist_from_zero, y = deployment_depth, col = ' '), pch = 23, size = 3, fill = 'grey') +
  scale_color_manual(values = c(" " = "#505050")) +
  scale_y_reverse(expand = c(0, 0)) +
  scale_x_continuous(expand = c(0, 0)) +
  xlab("Distance from western shore (m)") +
  ylab("Depth (m)") +
  labs(fill = expression('Temperature'~degree*C),
       col = "Loggers")

gp_lake_side_view

```


### Map of Europe
```{r}
worldMap <- rworldmap::getMap(resolution = "less islands")

# Member States of the European Union
europeanUnion <- c("Austria","Belgium","Bulgaria","Croatia","Cyprus",
                   "Czech Rep.","Denmark","Estonia","Finland","France",
                   "Germany","Greece","Hungary","Ireland","Italy","Latvia",
                   "Lithuania","Luxembourg","Malta","Netherlands","Poland",
                   "Portugal","Romania","Slovakia","Slovenia","Spain",
                   "Sweden","United Kingdom", "Switzerland")
# Select only the index of states member of the E.U.
EU <- worldMap[which(worldMap$NAME %in% europeanUnion),]
EU_sf <- st_transform(st_as_sf(EU), crs = CRS("+init=epsg:32633"))


gp_map <- ggplot() +
  geom_sf(data = EU_sf) +
  geom_sf(data = EU_sf %>% filter(NAME == "Czech Rep."), fill = "red")
  
gp_map

```

## Thermocline detection

### Temperature interpolation
First, each temperature profile was interpolated vertically (in depth) using monotone Hermite spline.
```{r}
ts_example <- "2015-09-06 10:00:00"

example_interval <- temperatures %>%
  filter(ts == ts_example)

hobo_example_interval <- hobo_data %>% filter(interval == ts_example)

example_interval_loggers <- inner_join(hobo_example_interval,
                                  hobo_deployments,
                                  by = c("location", "depth" = "deployment_depth"))

ggplot() +
  geom_line(data = example_interval, aes(x = depth, temperature, col = location)) +
  geom_point(data = hobo_example_interval, mapping = aes(x = depth, temperature, col = location))
```

## Thermocline interpolation over time
`rLakeAnalyzer` package was used to detect thermocline base, center and top in each profile (5min iterval at each locations). The method is good but it does not perform perfectly especially in case of weird profiles (e.g. multiple thermoclines, very mixed profiles, ...). Therefore, temperature of detected thermocline was smoothed by running median over 3 days for each location. Depthx of occurences of these smoothed temperatures were detected and finally smoothed by Fast Fourier Transform.

## Seasonal thermocline
```{r}
thermocline <- read_csv(here("data/products/thermocline_data.csv")) %>%
  filter(therm_part == "crit")

ggplot(data = thermocline %>% filter(therm_part == c("crit")), 
       aes(x = thermocline_ts, col = interaction(location))) + 
  geom_line(aes(y = depth)) +
  scale_y_reverse() +
  labs(col = "Location", x = "Time", y = "Depth (m)")
```

Seasonal thermocline depth was obtained for each location by fitting GAM model on computed thermocline depths.


#### Smoothing by Fast Fourier Transform
Smoothing was done on detrended data with influence of seasonal thremocline removed. The picture shows residual frequencies present in the time series.

```{r}
thermocline_fft <- read_csv(here("data/products/thermocline_fft.csv")) %>%
  filter(therm_part == "crit") %>%
  group_by(location, therm_part) %>% 
  mutate(x = row_number()) %>% 
  mutate(fft_spec_smoothed = ksmooth(x = x, y = abs(fft_spec), kernel = "box", bandwidth = 10)$y) %>%
  mutate(filtered_out = x < 700) %>%
  mutate(freq_hour = n() / (x * 12)) %>%
  ungroup()

n <- max(thermocline_fft$x)

gp_fs_hours <- c(24 * c(5, 2, 1),  c(12, 9, 7:2), 1, 0.5)
gp_breaks <- n / (gp_fs_hours * 12)

ggplot(thermocline_fft, aes(x = log(x),
                            y = fft_spec_smoothed,
                            col = location,
                            alpha = filtered_out))+
  scale_x_continuous(breaks = log(gp_breaks),
                     labels = gp_fs_hours,
                     limits = range(log(gp_breaks))) +
  scale_alpha_manual(values = c("TRUE" = 1, "FALSE" = 0.2)) +
    geom_line(stat = "identity") +
  labs(col = "Location", alpha = "Freq. used")+
  xlab("Frequency (hours)") +
  ylab("Spectral density")
```


##### Original data vs smoohed value
Snippet of final result of temporal interpolation of thermocline.

```{r}
date_range <- c("2015-07-06", "2015-07-13")
ggplot(data = thermocline %>% filter(therm_part == c("crit") &
                                       thermocline_ts > date_range[1] &
                                       thermocline_ts < date_range[2]), 
       aes(x = thermocline_ts, col = interaction(location))) + 
  geom_line(aes(y = depth, alpha = "original")) +
  geom_line(aes(y = Re(depth_fft_smoothed), alpha = "fft_smoothed")) +
  scale_alpha_manual(values = c("original"= 0.4, "smoothed" = 1)) +
  scale_y_reverse() +
  labs(col = "Location", x = "Time", y = "Depth (m)")
```


#### Various thermocline properties
* Strength - difference of temperatures between top and base
* Thickness - difference of depths between top and base
* Mean gradient - mean temperature gradient between top and base (strength/thickness)
* Deviation - difference between actual and smoothed thermocline depth
* Seasonal depth - depth computed using GAM model
### Correlation between variables
```{r}
thermocline_wide <- thermocline %>%
  pivot_wider(id_cols = c("location", "thermocline_ts", "thickness", "strength", "mean_gradient"),
              names_from = "therm_part",
              values_from = c("depth", "temperature")) %>%
  rename(ts = thermocline_ts)

thermocline_wide %>%
  dplyr::select(depth_crit, temperature_crit, thickness, strength, mean_gradient) %>%
  corrr::correlate() %>%
  corrr::rearrange() %>% 
  corrr::rplot(print_cor = T)
```

It is clear that thickness, strength and thermocline temperature are highly correlated with thermocline depth. It might be better to use mean gradient as another good descriptor of the temperature profile. 

There are other options to consider. E.g. compute index of overall stratification of the lake (literature suggests simply standard deviation of all temperatures, eventually there are some functions in `rLakeAnalyzer`).
